---
title: "APM466 AS1"
author: "Zefan Huang"
date: "14/02/2022"
output: pdf_document
---

```{r}
library(jrvFinance)
library(stringr)
DATA <- read.csv(file ="C:/Users/zefan/OneDrive/Desktop/selected.csv")
```

#Compute Dirty Price
```{r}
#Dirty price = Accrued interest + Clean Price
#Accrued interest = n/365* Annual Coupon Rate = F * c/M * D/T
#n = days since last payment  Face value is 100 here.
Clean_price <- DATA[c("X2022.01.10","X2022.01.11","X2022.01.12", "X2022.01.13","X2022.01.14","X2022.01.17","X2022.01.18","X2022.01.19","X2022.01.20","X2022.01.21")]
Price_date <- c("2022/01/10","2022/01/11","2022/01/12","2022/01/13","2022/01/14","2022/01/17","2022/01/18",
                   "2022/01/19","2022/01/20","2022/01/21")
Last_payment_date <- c("2021/09/01","2021/08/01","2021/08/01","2021/08/01","2021/08/01","2021/09/01","2021/09/01","2021/09/01","2021/09/01","2021/09/01")

                  
Dirty_price <- matrix(nrow=10,ncol=10)
coupon_payment <- as.numeric(sub("%","",DATA$coupon))#face value is 100
for (i in c(1:10)){
  for (j in c(1:10)){
    days_diff <- as.Date(as.character(Price_date[j]), format="%Y/%m/%d")-
                  as.Date(Last_payment_date[i], format="%Y/%m/%d")
    Dirty_price[i,j]=  Clean_price[i,j] + coupon_payment[i]*days_diff/365
  }
}
```

.
#compute YTM
Linear interpolation technique is used for this question,which is accurate enough
I compute YTM by future cash flows using bond.yield function from library jrvFinance.

```{r}
#bond.yield function in library jrvFinance can do the job.Clean prices are used here.
YTM <-matrix(nrow=10,ncol=10)
for (i in 1:10){
  for (j in 1:10) {
    YTM[i,j]=bond.yield(settle=Price_date[j],mature=DATA$maturity.date[i],coupon_payment2[i],freq=2,price=Clean_price[i,j],comp.freq=Inf, redemption_value=100)
  }
}

x <- c(0.5,1,1.5,2,2.5,3,3.5,4,4.5,5)
plot(x,YTM[,1],type="l",col="1",xlab="Time to Maturity (years)",ylab="Yield to Maturity",main="5-Year YTM Curves",ylim = c(0.002, 0.017))
for (i in 2:10){
lines(x,YTM[,i],col=i)
}
legend("bottomright",legend=Price_date,cex=0.5,col=1:10,lty=1)
```

#compute Spot rate
Using bootstrapping to compute the spot rates.
Algorithm:
Step1:Repeat 10 times for 10 dates.
Step2:Repeat 10 times for 10 bonds.
In the loop: Step3:For maturity less that 6 months, $r(T) = -\frac{log(\frac{P}{N})}{T}$,Here P is dirty price.(First bond Here.)
             Step4:For maturity more than 6 months(i.e. other 9 bonds),for j-th bond,$P = \sum_{h=1}^{j-1}P_{h}e^{-r(t_{h-1})t_{h-1}}+P_{j}e^{-r(t_{j})t_{j}}$, $r(t_{j-1})$ is calculated by previous process.
             $r_{j}=--\frac{log(\frac{P-\sum_{h=1}^{j-1}P_{h}e^{-r(t_{h-1})t_{h-1}}}{N+couponpayment})}{T}$
```{r}
Spot_rate <- matrix(nrow=10,ncol=10)
for (i in 1:10){
  P_today <- Dirty_price[,i]
  Spot_rate[i,1] <- -2*log(P_today[1]/(coupon_payment[1] + 100))
    for(j in 2:10){
    sum <- 0
    for (k in 1:(j-1)){
      sum <- sum + coupon_payment[j] * exp(-Spot_rate[i,k]*(0.5*k))
    }
    Spot_rate[i,j] <- log((P_today[j]-sum)/(coupon_payment[j] + 100))/((-1/2)*j)
  }
}
plot(x,Spot_rate[1,],type="l",col="1",xlab="Time to Maturity (years)",ylab="Spot Rate",main="5-Year Spot Rate Curves",ylim = c(0.004,0.03))
for (i in 2:10){
lines(x,Spot_rate[i,],col=i)
}
legend("bottomright",legend=Price_date,cex=0.5,col=1:10,lty=1)
```

#Compute forward rate
Algorithm:
Step1:Repeat 10 times for 10 dates.
Step2:Repeat 4 times for 4 1yr-nyr forward rates(n=1,2,3,4).
In the loop: Step3:1yr-1yr is just the spot rate.
             Step4:1yr-nyr forward rate:$f_{1-n}=(\frac{r(t_{2n})n-r(t_{2})}{n-1})$ ($r(t_{2})$ represents 1yr spot rate)
             
             
```{r}
y <- c(1,2,3,4)
For_rate <-matrix(nrow=4,ncol=10)
for (i in 2:5){
  For_rate[i-1,] <- (Spot_rate[,i*2] * i - Spot_rate[,2]) / (i- 1)
}
plot(y,For_rate[,1],type="l",col="1",xlab="Time to Maturity (years)",ylab="Forward Rate",main="Forward Rate",ylim = c(0.004,0.04),xlim = c(1,4))
for (i in 2:10){
lines(y,For_rate[,i],col=i)
}
legend("bottomright",legend=Price_date,cex=0.5,col=1:10,lty=1)
```

#Two covariance matrices
```{r}
y1<-c()
y2<-c()
y3<-c()
y4<-c()
y5<-c()
for (i in 1:9){
  y1[i]=log(YTM[2,i+1]/YTM[2,i])
  y2[i]=log(YTM[4,i+1]/YTM[4,i])
  y3[i]=log(YTM[6,i+1]/YTM[6,i])
  y4[i]=log(YTM[8,i+1]/YTM[8,i])
  y5[i]=log(YTM[10,i+1]/YTM[10,i])
}
yield_matrix <- data.frame(y1,y2,y3,y4,y5) 
yieldcov<-cov(yield_matrix)

f1 <- c()
f2 <- c()
f3 <- c()
f4 <- c()
for(i in 1:9)
{
  f1[i]=log(For_rate[1,i+1]/For_rate[1,i])
  f2[i]=log(For_rate[2,i+1]/For_rate[2,i])
  f3[i]=log(For_rate[3,i+1]/For_rate[3,i])
  f4[i]=log(For_rate[4,i+1]/For_rate[4,i])
}
for_matrix <-data.frame(f1,f2,f3,f4)
forcov <-cov(for_matrix)
```


#eigenvalues eigenvectors
```{r}
eigeny<-eigen(yieldcov)
eigenf<-eigen(forcov)
eigeny
eigenf
```

